---
title: "rstatix"
author: "The REAL Group 5"
date: "4/9/2022"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "../Output/")
```

# Introduction

rstatix is a r package for basic statistical tests that is pipe-friendly to R and RStudio. It can be used to add statistical analysis to plots. Some of the functions include normality tests, correlation analysis, student t-tests, and ANOVA.

For this project, we are using the PalmerPenguin data set to demonstrate how to use the package.

-   install.packages("ggpubr")
-   install.packages("rstatix")

## Load libraries
```{r}
library(tidyverse)
library(rstatix)
library(here)
library(ggplot2)
library(palmerpenguins)
library(ggpubr)
```

## Summary Statistics
Summary statistics is gathered to summarize and acquire information from the given data set.

This function allows us to compute summary statistics for either one or more numeric variables.

```{r}
#Full summary statistics
fullsum_penguins <- penguins %>% 
  get_summary_stats(year)
fullsum_penguins

#Grouped data summary statistics ; showing only common summary
groupsum_penguins <- penguins %>% 
  group_by(bill_length_mm, bill_depth_mm) %>% 
  get_summary_stats(year, type = "common")
groupsum_penguins

#Robust summary statistics
robustsum_penguins <- penguins %>% 
  get_summary_stats(year, type = "robust")
robustsum_penguins

#Summary statistics of only the mean and sd
meansdsum_penguins <- penguins %>% 
  get_summary_stats(year, type = "mean_sd")
meansdsum_penguins
```

## Identify Outliers 
Allows one to identify data outliers of a single variable at a time

```{r}
#Outlier of palmer penguin data for bill length (mm) 
outliers1 <- penguins %>%
  group_by(species)%>%
  identify_outliers(bill_length_mm)
view(outliers1)

#Outlier of palmer penguin data for bill depth (mm)
outliers2 <- penguins %>%
  group_by(species)%>%
  identify_outliers(bill_depth_mm)
view(outliers2)
```

## Shapiro Test
Single and multi-variable Shapiro-Wilk normality test.
This test examines if a variable is normally distributed in a population. 
Null hypothesis means data is evenly distributed.
p-value less than 0.05 rejects null hypothesis.
p-value greater than 0.05 does not reject null hypothesis. 

```{r}
#Shapiro test function
normality <- penguins %>%
  group_by(species) %>%
  shapiro_test(bill_length_mm, bill_depth_mm, body_mass_g, flipper_length_mm)
view(normality)

```

## t-test
t-tests are typically used to determine significant differences (p value) between the means of two group.

The rstatix package offers the t_test() function that can compute one-sample, two-samples, and pairwise t-tests. 

```{r}
## one-sample t-test
t_test1 <- penguins %>% 
  t_test(bill_length_mm ~ 1, mu = 0)
view(t_test1)

## two-samples paired t_test
t_test2 <- penguins %>% 
  t_test(bill_length_mm ~ sex) 
view(t_test2)

## How would you compute an unpaired t-test? Discuss with your neighbor
```

### t-test plotting
The p-value from t-tests can be used to plot via the function stat_pvalue_manual() with the ggpubr package.
We will be using the two-samples paired t-test from above to plot. 

```{r, bill_t_test, outwidth = "50%", fig.align='center'}
## First we need to install load the ggpubr package
## Next step is to remove any unwanted NAs from data and plot, so we need to have a clean data set
penguins_clean1 <- penguins %>% 
  drop_na(sex)  ### dropping NAs from the sex column ###
view(penguins_clean1)

## t-test plot with brackets
ggboxplot(penguins_clean1, x = "sex", y = "bill_length_mm", 
          color = "sex", palette = "simpsons", 
          facet.by = "year", ylim = c(0,70)) +
  stat_pvalue_manual(t_test2, label = "p", y.position = 65) + 
  labs(title = "Penguin Bill Length T-test", 
       x = "Sex", y = "Bill Length (mm)", 
       caption = "Source: Palmerpenguins package")
  ### to remove brackets, use "remove.bracket = TRUE" within the stat_pvalue_manual function

## t-test plot that shows significance 
ggboxplot(penguins_clean1, x = "sex", y = "bill_length_mm", 
          color = "sex", palette = "simpsons", 
          facet.by = "year", ylim = c(0,70)) +
  stat_pvalue_manual(t_test2, label = "p.adj.signif", y.position = 65) +
  labs(title = "Penguin Bill Length T-test", 
       x = "Sex", y = "Bill Length (mm)", 
       caption = "Source: Palmerpenguins package")
  ### changes label from "p" to "p.adj.signif" ###
```

### Effect size for t-test
Cohen's d measure of effect size allows us to compute the effect size for a given t-test.

This is calculated as the difference between mean minus mu divided by the standard deviation.

Effect Sizes:

- 0.2: small effect
- 0.5: moderate effect
- 0.8: large effect
```{r}
penguin_clean <- penguins %>% 
  drop_na()

#t-test effect size for one sample
effect_one <- penguins %>% 
  cohens_d(year ~ 1, mu = 0)
effect_one

#T-test effect size for two independent samples
effect_two <-penguin_clean %>% 
  cohens_d(bill_length_mm ~ sex, var.equal = TRUE) 
effect_two
```


## ANOVA test
ANOVA testing (Analysis of Variance) is used to compare means between two or more independent groups. 
If there is an experiment that has one factor, the ANOVA test is called a **one-way ANOVA**. If there are two factors, then it is called a **two-way ANOVA**. A one-way ANOVA is a common from of testing using one grouping variable. 

This package has many different ANOVA tests that can be piped into the code:

-   Independent measures ANOVA
-   Repeated measures ANOVA
-   Mixed ANOVA
-   ANCOVA: Analysis of Covariance

The table consists of several different columns:

-   DFn: degrees of freedom in the numerator
-   DFd: degrees of freedom in the denominator
-   F: F distribution (F-test)
-   p: p-value
-   p< 0.05: astrisks if the p-value is less than 0.05
-   ges: generalized effect size

```{r}
#for ANOVA testing, it is best to not have any NAs in the data
penguins_clean <- penguins %>%
  drop_na() 

#one-way ANOVA
#variable: bill length
#can get anova table in the console by using get_anova_table
#objecting the test helps to open the anova table in a new tab
one_way_pens <-penguins_clean %>%
  anova_test(year ~ bill_length_mm)
get_anova_table(one_way_pens, correction = "none")


#grouped one-way ANOVA
#grouping by species
group_anova <- penguins_clean %>%
  group_by(species) %>%
  anova_test(year ~ bill_length_mm)
get_anova_table(group_anova)


#two-way ANOVA
#two variables: bill length and depth
two_way_pens <- penguins_clean %>%
  anova_test(year ~ bill_length_mm*bill_depth_mm)
get_anova_table(two_way_pens, correction = "none")


#grouped two-way ANOVA
group_twopens <- penguins_clean %>%
  group_by(species) %>%
  anova_test(year ~ bill_length_mm*bill_depth_mm)
get_anova_table(group_twopens)
```

### ANOVA test plotting

```{r, anova_plot, outwidth = "50%", fig.align='center'}
ggplot() +
  geom_boxplot(penguins_clean, mapping = aes(x = year, y = bill_length_mm, fill = island )) +
  facet_wrap(~species) +
  labs(title = "Penguin Bill Length Across Species",
       x = "Year",
       y = "Bill Length (mm)",
       fill = "Island",
       caption = "Source: PalmerPenguins")
```