---
title: "rstatix_proj"
author: "The REAL Group 5"
date: "4/9/2022"
output: 
  html_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

rstatix is a r package for basic statistical tests that is pipe-friendly to R and RStudio. It can be used to add statistical analysis to plots. Some of the functions include normality tests, correlation analysis, student t-tests, and ANOVA.

For this project, we are using the PalmerPenguin data set to demonstrate how to use the package.


## Load libraries
```{r}
library(tidyverse)
library(rstatix)
library(here)
library(ggplot2)
library(palmerpenguins)
```

## First code function

```{r}

```

## Second code function

```{r}

```

## Third code function

```{r}

```

## T-test

```{r}

```

### T-test plotting

```{r}

```

## ANOVA test
ANOVA testing (Analysis of Variance) is used to compare means between two or more independent groups. 
If there is an experiment that has one factor, the ANOVA test is called a **one-way ANOVA**. If there are two factors, then it is called a **two-way ANOVA**. A one-way ANOVA is a common from of testing using one grouping variable. 

This package has many different ANOVA tests that can be piped into the code:

-   Independent measures ANOVA
-   Repeated measures ANOVA
-   Mixed ANOVA
-   ANCOVA: Analysis of Covariance

```{r}
penguins_clean <- penguins %>%
  drop_na() #for ANOVA testing, it is best to not have any NAs in the data

#one-way ANOVA
one_way_pens <-penguins_clean %>%
  anova_test(year ~ bill_length_mm)
get_anova_table(one_way_pens)

#grouped one-way ANOVA
group_anova <- penguins_clean %>%
  group_by(species) %>%
  anova_test(year ~ bill_length_mm)
get_anova_table(group_anova)

#two-way ANOVA
two_way_pens <- penguins_clean %>%
  anova_test(year ~ bill_length_mm*bill_depth_mm)
get_anova_table(two_way_pens)

#grouped two-way ANOVA
group_twopens <- penguins_clean %>%
  group_by(species) %>%
  anova_test(year ~ bill_length_mm*bill_depth_mm)
get_anova_table(group_twopens)
```

### ANOVA test plotting

```{r}

```